Apache Storm

### 2018-01-31 paper.tex
Begin changes as per suggestion

\textbf{Data} today is generated by a lot of devices that are connected to the internet and the amount of data being generated at any instance is increasing rapidly. These devices could include a personal computer, a smart phone, or even an embedded device. Due to the rapid adoption of Internet of Things, data is being generated at an unprecedented rate. The sheer volume of data generated and the rate at which it is being generated can be termed as Big Data. \textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.\cite{miningstatus}

@@

@@

Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the most easy to use, fast and fault tolerant systems available \cite{stormperspective}. These three factors have been the major points for the general public to choose \textbf{Apache Storm} above other available tools. It was originally created by Nathan Marz at BackType and till now, at present more than 60 companies use or experiment with this tool.

\subsection{Architecture Overview}
The basic Storm data processing architecture consists of stream of tuples flowing through a defined topology to yield the required analytic result. Processing of this stream data is performed in Nodes called the worker nodes, which are responsible for performing the required calculations. There is also another node named the Master Node, which is responsible for the complete Storm infrastructure,  by keeping track of all the worker nodes and the total topology itself.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{arch.png}
    \caption{High Level Storm Architecture}
    \label{fig:arch}
   \end{center}
\end{figure}

\subsubsection{Nimbus}

\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for the distribution and coordination of the execution of topologies. The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.  Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states. The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology \cite{stormtwitter}. It is also worth noting that multiple worker processes in a worker node could be executing different parts of the topology.

\subsubsection{Worker Node}
\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the topology. One worker node may run a multiple number of worker processes. Each of these worker processes may belong to the same topology or different topologies. A worker node maintains a \textbf{Supervisor} and a multiple number of Worker Processes which in turn maintains a multiple number of executors. Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism. Figure \ref{fig:workerarch} depicts the design of a Worker node.\cite{stormtwitter}

@@

\paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it. Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the worker processes and respawns the worker processes in case it dies.

\paragraph{Worker Process} is the Entity that does the processing and is mapped to a certain topology. Each worker node spawns a JVM and runs one or multiple executors. These executors may be running any of the bolt or spout within the topology, thus providing an added level of parallelism.

\paragraph{Executors} are made up of one or multiple Tasks. Unless specified explicitly, storm assigns single task to a single executor. Executors take tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.

\subsection{Topology}
\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represent the data flow. Another important thing about topologies in Storm is that they are allowed to have cycles. Figure \ref{fig:topo} shows a basic Storm Topology, which consists of Spouts and Bolts. Spouts are the Sources of tuples in a Topology whereas Bolts are the units which process these generated tuples. The user first submits the topology to the Apache Storm Nimbus, which consults with each of the available Supervisors in the worker nodes and assigns each of them with tasks related to the topology.

\subsubsection{Spout} 
\paragraph{Spouts} are the entry point for any data in the Storm Topology. As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples. The input source could be a Heterogenous data source.

\subsubsection{Bolt} 
\paragraph{Bolts} are the processing entities of a Topology. It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.

@@

\subsection{Fault Tolerance and Resilience}
\label{sec:resilience}
Apache storm offers fault tolerance and resilience mechanisms with out of the box support and is tolerant against different types of failures. Nimbus, being the main master node, is in-charge of all the other worker nodes. Under the circumstance that a worker node fails, Nimbus re-spawns a worker node and assigns the tasks to this node. If the nimbus itself fails, the nimbus is not re-started automatically, but the worker nodes do not stop processing the stream. The nimbus can be later on restarted manually. During this phase, if a worker fails, it will not be restarted. The supervisor is also an active element within the Storm Architecture and keeps the Nimbus posted about it's current status. The \textbf{heartbeat} event is scheduled to run every 15 seconds and it reports to the Nimbus that the worker node is alive. \cite{stormtwitter}

@@

\paragraph{Solution} to the scenario can be implemented using \textbf{Apache Storm} to process the tweets in real time. Figure \ref{fig:solution} shows the simplified system overview, consisting of Twitter Streaming Servers as the source of information. This information is processed in the Storm Clusters from within the Company Infrastructure and the analyzed result is reported to the User Interface, which can be accessed by the appropriate user.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{solution.png}
    \caption{Generic Streaming Solution implemented in this use case}
    \label{fig:solution}
   \end{center}
\end{figure}

>>>>

\textbf{Data} today is generated by a lot of devices that are connected to the internet and the amount of data being generated at any instance is increasing rapidly. These devices could include a personal computer, a smart phone, or even an embedded device. Due to the rapid adoption of Internet of Things, data is being generated at an unprecedented rate. The sheer volume of data generated and the rate at which it is being generated can be termed as Big Data. The word \textbf{Big Data} made it's appearance in 1998 SGI slide deck made by John Mashey and now is of interest to a lot of financial as well as research institutions.\cite{miningstatus}

@@

\section{Stream Processing}
A stream is a continuous flow of data from one point to the other. Stream Processing System works on these data streams, the task of these processing scheme may range from simple transformations to something as complex as event processing. In this present digitized world, people tend to post most of their thoughts and voices about topics on the web. These may be in the form of a long text, generally a webpage called blogs, or as a short post on a social networking site. An example of a social networking site is Twitter. In twitter, people tend to post short views about various things and at any moment there are a number of people posting onto twitter.

A company which is interested in analyzing the current outlook of the organization can decide to analyze these posts on twitter. The posts created by twitter users could now then be considered as Data Stream for the company. As seen in \ref{fig:example} various users generate posts on twitter, which is stored on the twitter servers. These tweets are used by the company to analyze the sentiment of the users and the Reporting UI represents the results to the respective interested party.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{example.png}
    \caption{An Example Stream Processing Scenario}
    \label{fig:example}
   \end{center}
\end{figure}

\subsection{Stream Processing Models}
There are two types of stream processing models, namely continuous flow model and discretized stream model. \cite{streaming}

\subsubsection{Continuous Stream Processing}

\subsubsection{Discretized Stream Processing}

@@

Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the simplest, fastest and fault tolerant systems available \cite{stormperspective}. These three factors have been the major points for the general public to choose \textbf{Apache Storm} above other available tools. It was originally created by Nathan Marz at BackType and till now, at present more than 60 companies \cite{stormtwitter} use or experiment with this tool.

\subsection{Architecture Overview}
The physical high level architecture of Storm as seen in figure \ref{fig:physicalarc} consists of a single master node and multiple worker nodes. The master node(see later) is the leader of the cluster and is responsible in short for the management of the different parts of the topology. There can only be one master node in a cluster. The worker node on the other hand is responsible for the actual execution of tasks supplied by the user. A user task is defined in storm as a Topology which actually is a directed graph. The vertices of the graph represent the data processing being done and the edges the flow of data. A data stream passed through the topology yields the required processing result as defined by the user.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{arch.png}
    \caption{High Level Storm Physical Architecture}
    \label{fig:physicalarc}
   \end{center}
\end{figure}

The logical or the runtime architecture of storm is defined within a topology. As seen in figure \ref{fig:logicalarch} shows an example of a topology and consists of vertices and edges, the vertices represent point where either data is generated or some processing to the existing data is done. The edges represent the flow of data, and in a topology, the data is represented in the form of tuples. A vertex can in storm be either a spout or a bolt. A spout is the source of tuples and it injects tuples into the topology, wheras a bolt is a vertice within a topology where the injected tuples are processed.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{arch.png}
    \caption{High Level Storm run-time or Logical Architecture}
    \label{fig:logicalarch}
   \end{center}
\end{figure}

The logical and physical architectures meet while being deployed in a cluster, the worker nodes are responsible for executing the topology. Every worker spawns a number of processes which run topologies. See later for detailed explanation on how topologies can be divided and executed within a storm cluster.

\subsubsection{Storm Components}
\paragraph{Master node} also called the \textbf{Nimbus} is responsible for the distribution and coordination of the execution of topologies. The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes. The user submits the code as a jar file to the nimbus, which in turn uses a combination of local disk and the zookeeper to store the topology defined in the jar file.\cite{stormtwitter} The topology thrift objects are stored in the zookeeper, whereas the code itself is stored in the local disk of nimbus.
\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the incoming data stream. A worker node maintains a \textbf{Supervisor} process and multiple worker processes. The number of  worker processes depends upon the storm configurations. A worker process once spawned, starts working for a single topology alone and not for a multiple number of topologies. It creates threads called \textbf{Executors} which take up a tasks to be performed. A task for an executor is the vertex of the Directed graph, ie. either a bolt or a spout. This can be observed in the \ref{fig:workerarch}, where a worker node has multiple worker processes, which in turn has multiple executors running different tasks. 
Even though storm can run multiple topologies, different processes run separate topologies.

The introduction of tasks enables a parallel execution of a number of same bolt or spout node of a topology, whereas the introduction of executors enables us to run different parts of the topology namely bolts and spouts in parallel.\cite{stormtwitter}. It is worth noting that these different parts of the topology, namely bolts and spouts can be executing within a single worker process in different executors, or can also be in different executors across different worker nodes.

@@

\paragraph{Supervisor} is part of a Worker Node and communicates with Nimbus and accepts incoming jobs. It monitors the health of the worker processes and respawns every process in case of failure. The supervisor performs three major periodic tasks, namely heartbeat, synchronize supervisor and sunchronize process.
The \textbf{heartbeat} event is scheduled every 15 seconds by the supervisor and in this event it reports to the nimbus that the worker node is alive and provides information about the available worker processes.\cite{stormtwitter} With this provided information, nimbus decides whether it can accept new jobs or not and depending upon the requirements, schedules jobs to be performed by the node. The \textbf{synchronize supervisor} event runs every 10 seconds and is reponsible for synchronization of the supervisor. It monitors the change is existing assignments and if the change includes an addition of topology, then downloads the required jar files and its dependencies and immediately schedules a synchronize worker process event.

The \textbf{synchronize worker process} event runs every 3 seconds and is responsible for managing worker processes. It reads worker heartbeats and classifies those workers with one of the labels, \textit{valid, timed out, not started or disallowed}.

\subsubsection{Topology}
\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represent the data flow. Another important thing about topologies in Storm is that they are allowed to have cycles. Figure \ref{fig:topo} shows a basic Storm Topology, which consists of Spouts and Bolts. Spouts are the Sources of tuples in a Topology whereas Bolts are the units which process these generated tuples. The user first submits the topology to the Apache Storm Nimbus, which consults with each of the available Supervisors in the worker nodes and assigns each of them with tasks related to the topology.

\paragraph{Spouts} are the entry point for any data in the Storm Topology. As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples. The input source could be a Heterogenous data source.

\paragraph{Bolts} are the processing entities of a Topology. It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.

@@

\subsection{Storm Features}
\subsubsection{Fault Tolerance and Resilience}
\label{sec:resilience}
Apache storm offers fault tolerance and resilience mechanisms with out of the box support and is tolerant against different types of failures. Nimbus, being the main master node, is in-charge of all the other worker nodes. Under the circumstance that a worker node fails, Nimbus re-spawns a worker node and assigns the tasks to this node. If the nimbus itself fails, the nimbus is not re-started automatically, but the worker nodes do not stop processing the stream. The nimbus can be later on restarted manually. During this phase, if a worker fails, it will not be restarted. The supervisor is also an active element within the Storm Architecture and keeps the Nimbus posted about it's current status. The \textbf{heartbeat} event is scheduled to run every 15 seconds and it reports to the Nimbus that the worker node is alive. \cite{stormtwitter}

@@

\paragraph{Solution} to the scenario can be implemented using \textbf{Apache Storm} to process the tweets in real time. Figure \ref{fig:example} shows the simplified system overview, consisting of Twitter Streaming Servers as the source of information. This information is processed in the Storm Clusters from within the Company Infrastructure and the analyzed result is reported to the User Interface, which can be accessed by the appropriate user.
